---
title: "Numeric Clustering"
author: "Laurae"
date: "October 06, 2016"
output: html_notebook
---

# Introduction

This prints the Empirical Cumulative Density Function (ECDF) of all numeric features.

# Plot interpretation

The gap of Response ECDF at the furthest right correspond to the NAs.

The plots are titled the following:

* Title: ECDF of Response vs Feature
* Subtitle: NA = amount of NAs (Missing: % of missing -> Response: % of responses accounting to missing)

The X axis depicts the numeric feature assessed, while the Y axis is for the ECDF of Response vs the numeric feature.

# Setting up work

```{r}
# Load libraries
library(data.table)
library(recommenderlab)
library(pbapply)

# Set your working directory there =)
setwd("E:/")

# Load data
data <- readRDS("datasets/train_numeric.rds")
gc(verbose = TRUE)
```

# Plotting madness =)

```{r}
for (i in 1:969) {
  
  cat("[", format(Sys.time(), "%a %b %d %Y %X"), "] Doing feature ", sprintf("%03d", i), "...  \n", sep = "")
  #ecdfed <- ecdf(data[[i]])
  #plot(ecdfed)
  temp_NA <- sum(is.na(data[[i]]))
  if ((min(data[[i]], na.rm = TRUE) == max(data[[i]], na.rm = TRUE)) | (sum(is.na(data[[i]])) == 1183747)) {
    # do nothing
  } else {
    plot(sort(data[[i]], na.last = TRUE), (1:1183747)/1183747,
         type = "s",
         ylim = c(0, 1),
         main = paste("Empirical Cumulative Density Function (ECDF) of Response vs ", colnames(data)[i], "\nNA = ", temp_NA, " (Missing: ", sprintf("%05.2f", 100 * temp_NA / 1183747), "% -> Response: ", sprintf("%05.2f", 100 * sum(data$Response[is.na(data[[i]])]) / 6879), "%)", sep = ""),
         xlab = colnames(data)[i],
         ylab = "f(Response)"
    )
  }
  gc(verbose = FALSE)
  
}
```

