---
title: "Numeric (Conditional) Mutual Information"
output:
  html_notebook:
    css: discretization.css
---

# Preload

Preload things.

```{r}
setwd("E:/")
library(recommenderlab)
library(data.table)
library(infotheo)
library(DT)
train <- readRDS("datasets/train_numeric.rds")
date <- readRDS("Laurae/DateFeaturesMore/Min_features.rds")
#train <- readRDS("Laurae/NumericCMI_exact/train_numeric_discrete.rds")
#date <- readRDS("Laurae/NumericCMI_exact/train_date_discrete.rds")
feature_names <- fread("Laurae/date_feat_lut_V2.txt", header = TRUE)
feature_names <- feature_names[!name_num == "", ]
feature_names <- feature_names[, c("name_num", "station_V2", "line", "station", "feature_nr"), with = FALSE]
colnames(feature_names) <- c("Feature", "Station", "Line", "Orig_St", "Feat_Nb")
gc(verbose = TRUE)
knitr::opts_chunk$set(root.dir = "E:/") # not working??! 
```

# Information Gain

## Preprocessing by complete discretization (hard-level)

```{r}
ig_frame1 <- data.frame(matrix(nrow = 968, ncol = 7))
colnames(ig_frame1) <- c("Feature", "Levels", "IG_Numeric", "IG_Date", "CMI", "NotMissing", "Number")
ig_frame1$Feature <- colnames(train)[2:969]
ig_frame1$Number <- 1:968
pb <- winProgressBar(title = "Discretization of Numerics", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Preparing computation...", sep = ""), min = 0, max = 968, initial = 0, width = 520)
for (i in 1:968) {
  my_na <- is.na(train[[i + 1]])
  ig_frame1$NotMissing[i] <- sum(!my_na) / 1183747
  train[[i + 1]][my_na] <- -9999
  train[[i + 1]] <- as.numeric(as.factor(train[[i + 1]]))
  ig_frame1$Levels[i] <- length(unique(train[[i + 1]]))
  gc(verbose = FALSE)
  setWinProgressBar(pb, value = i, title = "Discretization of Numerics", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Doing feature ", colnames(train)[i + 1], " (", sprintf("%03d", i), " / 968 = ", sprintf("%05.2f", 100 * i / 968), "%)...", sep = ""))
}
close(pb)
saveRDS(train, "E:/Laurae/NumericCMI_exact/train_numeric_discrete.RDS", compress = TRUE)
```

```{r}
ig_frame2 <- data.frame(matrix(nrow = 128, ncol = 5))
colnames(ig_frame2) <- c("Feature", "Levels", "IG_Date", "NotMissing", "Number")
ig_frame2$Feature <- colnames(date)
ig_frame2$Number <- 1:128
date <- date[1:1183747, ]
pb <- winProgressBar(title = "Discretization of Dates", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Preparing computation...", sep = ""), min = 0, max = 128, initial = 0, width = 520)
for (i in 1:128) {
  my_na <- is.na(date[[i]])
  ig_frame2$NotMissing[i] <- sum(!my_na) / 1183747
  date[[i]][my_na] <- -9999
  date[[i]] <- as.numeric(as.factor(date[[i]]))
  ig_frame2$Levels[i] <- length(unique(date[[i]]))
  gc(verbose = FALSE)
  setWinProgressBar(pb, value = i, title = "Discretization of Dates", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Doing feature ", colnames(date)[i], " (", sprintf("%03d", i), " / 128 = ", sprintf("%05.2f", 100 * i / 128), "%)...", sep = ""))
}
close(pb)
saveRDS(date, "E:/Laurae/NumericCMI_exact/train_date_discrete.RDS", compress = TRUE)
```
## Precompute all information gain values

Formulas:

Mutual Information = I(X, Y) = H(X, Y) - H(X | Y) - H(Y | X), understand un-sided
Conditional Mutual Information = I(X, Y | Z) = H(Z | X) + H(X) + H(Z | Y) + H(Y) - H(Z | X, Y) - H(X, Y) - H(Z) = I(X, Y) + H(Z | X) - H(Z | Y) - H(Z | X, Y) - H(Z), with guarantee of I(X, Y | Z) > 0 and being one-sided

```{r}
pb <- winProgressBar(title = "Date Mutual Information computation", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Preparing computation...", sep = ""), min = 0, max = 128, initial = 0, width = 520)
for (i in 1:128) {
  setWinProgressBar(pb, value = i, title = "Date Mutual Information computation", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Doing feature ", colnames(date)[i], " (", sprintf("%03d", i), " / 128 = ", sprintf("%05.2f", 100 * i / 128), "%)...", sep = ""))
  ig_frame2$IG_Date[i] <- condinformation(X = date[[colnames(date)[i]]], Y = train$Response, S = NULL, method = "emp")
  gc(verbose = FALSE)
}
close(pb)
```

```{r}
pb <- winProgressBar(title = "Numeric Mutual Information computation", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Preparing computation...", sep = ""), min = 0, max = 968, initial = 0, width = 520)
for (i in 1:968) {
  setWinProgressBar(pb, value = i, title = "Numeric Mutual Information computation", label = paste("[", format(Sys.time(), "%a %b %d %Y %X"), "] Doing feature ", colnames(train)[i + 1], " (", sprintf("%03d", i), " / 968 = ", sprintf("%05.2f", 100 * i / 968), "%)...", sep = ""))
  my_feature <- colnames(train)[i + 1]
  my_date <- paste("S", feature_names$Station[which(my_feature == feature_names$Feature)], "_Min", sep = "")
  ig_frame1$IG_Numeric[i] <- condinformation(X = train[[my_feature]], Y = train$Response, S = NULL, method = "emp")
  ig_frame1$IG_Date[i] <- ig_frame2$IG_Date[which(my_date == ig_frame2$Feature)]
  ig_frame1$CMI[i] <- condinformation(X = train[[my_feature]], Y = train$Response, S = date[[my_date]], method = "emp")
  gc(verbose = FALSE)
}
close(pb)
```

## Setup feature names properly for Numeric

```{r}
all_frame <- merge(ig_frame1, feature_names, by = "Feature", all.x = TRUE)
all_frame <- all_frame[, c(1, 7, 9, 8, 10, 11, 2, 3, 4, 5, 6)]
all_frame <- all_frame[order(all_frame$Number), ]
rownames(all_frame) <- NULL
all_frame$Number <- NULL
all_frame$Num_Rank <- 969 - rank(all_frame$IG_Numeric, ties.method = "max")
all_frame$Date_Rank <- 969 - rank(all_frame$IG_Date, ties.method = "max")
all_frame$CMI_Rank <- 969 - rank(all_frame$CMI, ties.method = "max")
#all_frame$Relative_Gain <- 1 - ((max(all_frame$Information_Gain, na.rm = TRUE) - all_frame$Information_Gain) / (max(all_frame$Information_Gain, na.rm = TRUE) - min(all_frame$Information_Gain, na.rm = TRUE)))
gc(verbose = TRUE)
fwrite(all_frame, "E:/Laurae/NumericCMI_exact/CMI_numeric.csv")
```

## Pretty print DT for Numeric

```{r}
datatable(all_frame,
          filter = "top",
          class = "cell-border stripe",
          extensions = c("AutoFill",
                         #"Buttons",
                         "ColReorder",
                         "KeyTable",
                         "Responsive",
                         "RowReorder"),
          options = list(pageLength = 20,
                         lengthMenu = c(5, 10, 15, 20, 25, 50, 100, 500, 1000),
                         order = list(list(9, "desc")),
                         autofill = TRUE,
                         #dom = "Bfrtip",
                         #buttons = c("copy", "csv", "excel", "pdf", "print"),
                         colReorder = TRUE,
                         keys = TRUE,
                         rowReorder = TRUE)
          ) %>% formatStyle('IG_Numeric',
                             background = styleColorBar(range(all_frame$IG_Numeric, na.rm = TRUE, finite = TRUE), 'lightgreen'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatStyle('IG_Date',
                             background = styleColorBar(range(all_frame$IG_Date, na.rm = TRUE, finite = TRUE), 'lightblue'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatStyle('CMI',
                             background = styleColorBar(range(all_frame$CMI, na.rm = TRUE, finite = TRUE), 'pink'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatStyle('NotMissing',
                             background = styleColorBar(c(0, 1), 'lightgrey'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatStyle('Levels',
                             background = styleColorBar(range(all_frame$Levels, na.rm = TRUE, finite = TRUE), 'lightgrey'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatRound(columns = c("IG_Numeric"),
                            digits = 8) %>%
                formatRound(columns = c("IG_Date"),
                            digits = 8) %>%
                formatRound(columns = c("CMI"),
                            digits = 8) %>%
                formatPercentage(columns = c("NotMissing"),
                                 digits = 4)
```

## Setup feature names properly for Date

```{r}
all_frame <- merge(ig_frame2, feature_names, by = "Feature", all.x = TRUE)
all_frame <- all_frame[, c(1, 5, 2, 3, 4)]
all_frame <- all_frame[order(all_frame$Number), ]
rownames(all_frame) <- NULL
all_frame$Number <- NULL
#all_frame$Relative_Gain <- 1 - ((max(all_frame$Information_Gain, na.rm = TRUE) - all_frame$Information_Gain) / (max(all_frame$Information_Gain, na.rm = TRUE) - min(all_frame$Information_Gain, na.rm = TRUE)))
all_frame$IG_Rank <- 129 - rank(all_frame$IG_Date, ties.method = "max")
gc(verbose = TRUE)
fwrite(all_frame, "E:/Laurae/NumericCMI_exact/CMI_date.csv")
```

## Pretty print DT for Date

```{r}
datatable(all_frame,
          filter = "top",
          class = "cell-border stripe",
          extensions = c("AutoFill",
                         #"Buttons",
                         "ColReorder",
                         "KeyTable",
                         "Responsive",
                         "RowReorder"),
          options = list(pageLength = 20,
                         lengthMenu = c(5, 10, 15, 20, 25, 50, 100, 500),
                         order = list(list(3, "desc")),
                         autofill = TRUE,
                         #dom = "Bfrtip",
                         #buttons = c("copy", "csv", "excel", "pdf", "print"),
                         colReorder = TRUE,
                         keys = TRUE,
                         rowReorder = TRUE)
          ) %>% formatStyle('IG_Date',
                             background = styleColorBar(range(all_frame$IG_Date, na.rm = TRUE, finite = TRUE), 'lightgreen'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatStyle('NotMissing',
                             background = styleColorBar(c(0, 1), 'lightgrey'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatStyle('Levels',
                             background = styleColorBar(range(all_frame$Levels, na.rm = TRUE, finite = TRUE), 'lightgrey'),
                             backgroundSize = '100% 90%',
                             backgroundRepeat = 'no-repeat',
                             backgroundPosition = 'center') %>%
                formatRound(columns = c("IG_Date"),
                            digits = 8) %>%
                formatPercentage(columns = c("NotMissing"),
                                 digits = 4)
```

